In this paper, we gave a new method to compute some dynamical properties
on Process Hitting models, a subclass of asynchronous automata.
The main interest of our method is the use of ASP,
a declarative programming paradigm which benefits from powerful solvers.
We first focused on the enumeration of the fixed points of a model,
that is, the states from which no more evolution is possible,
which is tackled simply on Process Hitting models given their particular form.
We also considered the reachability problem, that is,
checking if it is possible to reach a state with a given property
from a given initial state,
which thus corresponds to an $\mathsf{EF}$ property in CTL logic.
Our analysis is exhaustive:
it computes the dynamics of the model from a given initial state and checks the sought goal;
this search can also be limited to a given number of steps.
%for which the dynamics of the model from the given initial state is computed.
%and can be limited to a number of steps,

We gave an implementation of these problems into ASP,
and applied them to several biological examples of various sizes, up to
40 biological components.
The results showed that our implementation is faster and deals with bigger models
than other classical approaches, especially \textsc{LibDDD} which is a symbolic model-checker.

Our work could benefit from several extensions.
Of course, the set of applicable models can be extended,
for example with the addition
of priorities or neutralizing edges,
or by considering synchronous dynamics or other representations
such as Thomas modeling~\cite{BernotSemBRN}.
However, the range of the analysis can also be extended,
by searching instead the set of initial states
allowing to reach a given goal,
or extending the method to universal properties (like the $\mathsf{AF}$ operator).
